---
title: "Notebook version of FluvialStratApp"
output: html_notebook
---
This is a simple forward stratigrahpic model for fluvial deposits. This is a notebook that has 5 code chunks that go from loading libraries, through the model, to plotting the model output in two different ways.


Required packages/libraries. Make sure that these libraries are installed in R Studio.
```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(gridExtra)
```

Set/define key variables.
```{r}
#Model/basin controls
model_width <- 10000 #width of modelling space (m) 
sub_rate <- 0.0025 #subsdience rate (m/yr)
strat_th <- 50 #thickness of stratigraphy to generate (m)

#Channel belt controls
blt_wd <- 2000 #channel belt width (m)
blt_p_avl <- 0.00001 #probability of channel belt avulsion, p=1/belt avulsion frequency 

#Channel form controls
ch_wd <- 150 #channel width (m)
ch_th <- 5 #channel thickness (m)
ch_tri <- 1 #triangular-ness of the channelform (ranging from 0-1, 0=rectangular channel deposit, 1=triangular channelform deposit)
rate <- 4 #migration rate in m/yr
ch_p_avl <- 0.0001 #probability of channel avulsion, p=1/channel avulsion frequency 
```

This is a genearl sanity check that merely displays the number of iterations
```{r}
print(paste("This will result in ", round(strat_th/sub_rate,digits=1), " model iterations",sep=""))
```

This is the whole model. It starts by loading a function that processes one channels worth of data. Then starts the whole model. It prints a plot fo the model at the end to display that it's done. It will also print updates as it goes on.
```{r}
#Load function for drawing channel polygons
draw_channel <- function(df){
    ch_df <- NULL #set up new object/dataframe for the new data
    local_mig_dir <- df[[1,11]] #get starting migration direction
    prev_ch_min <- df[[1,9]] #get starting ch_min
    prev_ch_max <- df[[1,10]] #get strating ch_max
    for (i in 1:nrow(df)){
        if((prev_ch_min+local_mig_dir)<df[[i,7]] | (prev_ch_max+local_mig_dir)>df[[i,8]]){ #if the channel is out of bounds,
            local_mig_dir <- (local_mig_dir*-1) #and flip the migration direction.
            x1 <- prev_ch_min + local_mig_dir #migrate it the other way,
            x4 <- prev_ch_max + local_mig_dir
        } else {
            local_mig_dir <- sample(c(local_mig_dir, (local_mig_dir*-1)),size=1,prob=c(0.8,0.2)) # give local_mig_dir a chance to switch 
            x1 <- prev_ch_min + local_mig_dir
            x4 <- prev_ch_max + local_mig_dir
        }
    prev_ch_min <- x1
    prev_ch_max <- x4
    
    ch_df <- rbind(ch_df,data.frame(x1=x1,
                                    y1=df[[i,1]],
                                    
                                    x2=x1+(ch_wd*(ch_tri/2)),
                                    y2=df[[i,1]]-ch_th,
                                    
                                    x3=x4-(ch_wd*(ch_tri/2)),
                                    y3=df[[i,1]]-ch_th,
                                    
                                    x4=x4,
                                    y4=df[[i,1]])
                   )
    
    } #end for loop
    processed_rows <<- as.numeric(processed_rows+(nrow(df)))
    print(paste("Processed ", processed_rows, "years (",(processed_rows/(years+1))*100,"% complete)",sep=""))
    cbind(df,ch_df)
}#end function

#------------------------

#Start of model
years <- strat_th/sub_rate
processed_rows <- 0
strat_df <- data.frame(thickness=seq(from=-strat_th,to=0, by=sub_rate),
                       blt_avl=sample(c(0,1),size=years+1,prob=c(1-blt_p_avl,blt_p_avl),replace=TRUE),
                       ch_avl=sample(c(0,1),size=years+1,prob=c(1-ch_p_avl,ch_p_avl),replace=TRUE)
            )  

strat_df[1,2:3] <- 1 #set the first row of the avulsion columns (blt_avl, ch_avl) to 1
strat_df[which(strat_df$blt_avl==1),3] <- 1 #set ch_avl==1 where blt_avl==1

strat_df$blt_no <- cumsum(strat_df$blt_avl) #create a belt number by cumulative sum
strat_df$ch_no <- cumsum(strat_df$ch_avl) #create channel number by cumulative sum
strat_df$dep_no <- paste(strat_df$blt_no,".",strat_df$ch_no,".",seq.int(nrow(strat_df)),sep="") #create unique deposit number by paste
print(paste("This model consists of: ", max(strat_df$blt_no), " channel belts and ", max(strat_df$ch_no), " channels", sep=""))

#mutate belt information
strat_df <- strat_df %>% 
    group_by(blt_no) %>% 
    mutate(blt_min=sample((-model_width/2):((model_width/2)-blt_wd), 1),
           blt_max=blt_min+blt_wd) %>% 
    ungroup()

#mutate channel information
strat_df <- strat_df %>% 
    group_by(ch_no) %>% 
    rowwise() %>%
    mutate(ch_min=sample((blt_min):(blt_max-ch_wd), 1),
           ch_max=(ch_min+ch_wd),
           mig_dir=sample(c(-rate,rate),size=1,prob=c(0.5,0.5))) %>%
    ungroup()
strat_df[which(strat_df$ch_avl==0),9:11] <- NA 

#split into a list of dataframes/tibbles
strat_list <- split(strat_df, f=strat_df$ch_no)

#lapply the draw_channel function to each dataframe/tibble in the list
strat_list <- lapply(strat_list, function(df)draw_channel(df))

#bind back into dataframe
strat_df <- do.call(rbind.data.frame, strat_list)

#filter into individual dataframes and rename x, y columns - I don't like pivoting... then bind into one data table
strata_x1_y1 <- strat_df %>% select(thickness,blt_no,ch_no,dep_no,blt_min,blt_max,x=x1,y=y1)
strata_x2_y2 <- strat_df %>% select(thickness,blt_no,ch_no,dep_no,blt_min,blt_max,x=x2,y=y2)
strata_x3_y3 <- strat_df %>% select(thickness,blt_no,ch_no,dep_no,blt_min,blt_max,x=x3,y=y3)
strata_x4_y4 <- strat_df %>% select(thickness,blt_no,ch_no,dep_no,blt_min,blt_max,x=x4,y=y4)

strat_df <- rbind(strata_x1_y1,strata_x2_y2,strata_x3_y3,strata_x4_y4)
belt_df <- strat_df %>% group_by(blt_no) %>% summarize(x_min=min(blt_min),
                                                       x_max=max(blt_max),
                                                       y_min=min(y),
                                                       y_max=max(y)
                                                       )

ggplot() +
    geom_rect(data=belt_df, mapping=aes(xmin=x_min,xmax=x_max,ymin=y_min,ymax=y_max,alpha=0.1)) +
    geom_polygon(data=strat_df, mapping=aes(x=x,y=y,fill=as.factor(ch_no),group=dep_no)) +
    coord_cartesian(xlim=c((-model_width/2),(model_width/2)),ylim=c(-strat_th,0)) + 
    labs(subtitle="Synthetic fluvial architecture", x="Distance (m)", y="Thickness (m)") +
    theme(legend.position = "none")
    

```


This is an example plotting for creating nieghbouring plots of the three different heirarchies.
```{r}
p <- ggplot() +
    geom_rect(data=belt_df, mapping=aes(xmin=x_min,xmax=x_max,ymin=y_min,ymax=y_max,alpha=0.05)) +
    coord_cartesian(xlim=c((-model_width/2),(model_width/2)),ylim=c(-strat_th,0)) + 
    labs(x="Distance (m)", y="Thickness (m)") +
    theme(legend.position = "none")

plot1 <- p + geom_polygon(data=strat_df, mapping=aes(x=x,y=y,fill=as.factor(blt_no),group=dep_no)) + labs(subtitle="Channel belts")
plot2 <- p + geom_polygon(data=strat_df, mapping=aes(x=x,y=y,fill=as.factor(ch_no),group=dep_no)) +  labs(subtitle="Channelform deposits")
plot3 <- p + geom_polygon(data=strat_df, mapping=aes(x=x,y=y,fill=as.factor(dep_no),group=dep_no)) + labs(subtitle="Individual deposits")

plot_heirarchy <- grid.arrange(plot1,plot2,plot3)
plot_heirarchy
```


Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
